---
title: "Homework 2, Alice"
output: html_document
---

Rename the blood_type_factor in the above example as follows:
levels(blood_type_factor) <- c("BT_A", "BT_B", "BT_AB", "BT_O")
and justify the result of outputting blood_type_factor. Use str() to support your answer.

```{r}
blood_type <- c("B", "A", "AB", "A", "O")
blood_type_factor <- factor(blood_type)
levels(blood_type_factor) <- c("BT_A", "BT_B", "BT_AB", "BT_O")
print(blood_type_factor)
str(blood_type_factor)
```
The issue seems to be order. When we make blood_type into a factor, it reorders things, but not in the order that I re-labeled the levels.

Use the inequality operator (< or >) on a nominal category and print the output.
```{r}
1 > 2
```

```{r}
blood_test <- list(Fluid="Blood sugar", Category="High", Value=140, Units="mg/dL")
patient <- list(Name="Mike", Age=36, Btest = blood_test)
```

What does patient[c(1,3)] give us? A list.
```{r}
patient[c(1,3)]
class(patient[c(1,3)])
```

What does patient[[c(1,3)]] give us? An error, because there's no 3rd item in the first item on the list.

```{r}
#patient[[c(1,3)]]
#Error in patient[[c(1, 3)]] : subscript out of bounds
```

How about patient[[c(3,1)]]? A character. It gave us the first value in the 3rd item.
```{r}
patient[[c(3,1)]]
class(patient[[c(3,1)]])
```

How do you remove an element from a list? I just removed the whole column, which probably isn't right. So for good measure, I also made the age into NA.
```{r}
patient$Name <- NULL
patient[c(1)] <- NA
patient
```

