---
title: "Homework 2, Eastoa01"
output: html_document
---

Rename the blood_type_factor in the above example as follows:  
levels(blood_type_factor) <- c("BT_A", "BT_B", "BT_AB", "BT_O")  
and justify the result of outputting blood_type_factor. Use str() to support your answer.

```{r}
blood_type <- c("B", "A", "AB", "A", "O")
blood_type_factor <- factor(blood_type)
levels(blood_type_factor) <- c("BT_A", "BT_B", "BT_AB", "BT_O")
print(blood_type_factor)
str(blood_type_factor)
```
The issue seems to be order. When we make blood_type into a factor, it reorders things, but not in the order that I re-labeled the levels.

Use the inequality operator (< or >) on a nominal category and print the output.
```{r}
1 > 2
```

```{r}
blood_test <- list(Fluid="Blood sugar", Category="High", Value=140, Units="mg/dL")
patient <- list(Name="Mike", Age=36, Btest = blood_test)
```

What does patient[c(1,3)] give us? *A list.*
```{r}
patient[c(1,3)]
class(patient[c(1,3)])
```

What does patient[[c(1,3)]] give us? *An error, because there's no 3rd item in the first item on the list.*

```{r}
#patient[[c(1,3)]]
#Error in patient[[c(1, 3)]] : subscript out of bounds
```

How about patient[[c(3,1)]]? *A character. It gave us the first value in the 3rd item.*
```{r}
patient[[c(3,1)]]
class(patient[[c(3,1)]])
```

How do you remove an element from a list? *I just removed the whole column, which probably isn't right. So for good measure, I also made the age into NA.*
```{r echo=TRUE}
patient$Name <- NULL
patient[c(1)] <- NA
patient
```

###Data frames
```{r include=FALSE}
name <- c("Anne", "James", "Mike", "Betty")
age <- c(20, 43, 27,25)
cancer <- c(TRUE, FALSE, FALSE, TRUE)
df <- data.frame(Name=name, Age=age, Cancer_Status=cancer)
df$Tumor_size <- c("T3", "T0", "T0", "T2")
tom <- data.frame(Name="Tom", Age=47, Cancer_Status="TRUE", Tumor_size="T2")
df <- rbind(df, tom)
```
My data frame is 4 observations of 3 variables.  Name is a factor with 4 levels, Age is a numeric variable, and Cancer_Status is logical.  
If you try to combine vectors with unequal length into a data frame, you get the following error message:

<span style="color:red">Error in data.frame(Name = name, Age = age, Cancer_Status = cancer) : arguments imply differing number of rows: 4, 3</span>

Test the operations of lists (like ["age"] & [["age"]]) on data frames.
```{r}
class(df["Age"]) # Prints a data frame with just one column
class(df[["Age"]]) #Prints a numeric vector
```
Use cbind() to append a vector of choice.  
```{r}
df2 <- cbind(df,Family_History=c(TRUE,TRUE, FALSE,FALSE))
```
What happens if the length of the appending vector is greater than (or less than) row dimensions? *If the appending vector has too many or too few elements, you get an error message about the number of rows.*

Can you use the list() function instead of the data frame function in the above code? *YES*
What happens if you leave the arguments name=, age= etc. in the above code?
```{r}
tom <- list(Name="Mary", Age=37, Cancer_Status="FALSE", Tumor_size="T0")
df <- rbind(df, tom)

```

What happens if you leave *out* the arguments name=, age= etc. in the above code? *It doesn't work; you get an error saying the names don't match.*

Check out the function expand.grid(), what is it for? *expand.grid creates a data frame from all combinations of factor variables.*
Try out the following example and think when it is useful.
```{r}
expand.grid(height = as.character(seq(60, 70, 5)), weight = seq(100, 200, 50),
            sex = c("Male","Female"), stringsAsFactors = FALSE)
```
*This creates
